A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-08-18 Papers/Paper/manuscript.Rmd
more edits

@@

Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT. The two most commonly used methods are eye-tracking and self-paced reading, but both of these suffer from a lack of localization. In eye-tracking, participants read a text on a screen naturally, while their eye-movements are recorded [@raynerEyeMovementsReading1998]. The downside of this natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982]. Self-paced reading is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead. Even in this method, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006]. Both these methods are prone to spillover, as the time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed, so by the time any late processing difficulties occur, the reader may already be several words along.

@@

## How to compare localization & effects, (old school freq/length), define surprisal and lang models
TODO!!!
One measure of how localized a method is is to look at how much overall effects on each word stay on that word, versus appearing on downstream words. 
There are a few general effects for how properties of words tend to effect their reading time. 
The classic pair are frequency and length -- longer and less common words take longer to process. @klieglLengthFrequencyPredictability2004
Another measure is predictability, or how expected a word is in context, with more predictable words being read faster @raynerEffectsFrequencyPredictability2004. A word can have a high surprisal for lots of reasons -- it could be low frequency, or semantically unexpected, or the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse. Many targeted effects of interest contribute to how predictable or unpredictable a word is. If effects of predictability are strong and well-localized, it's a good sign. 
Because predictability is made up of all these components, how do we assess how predictable a word is? One way of doing this is used language models that are trained to predict the next word. Predictability is often measured in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc). 
RTs for eye-tracking and SPR show a linear relationship between a words surprisal and RT It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
 Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.

## why current experiment
TODO!!!
Maze is great, but there hasn't been a study looking at these coarse effects. We're interested in whether there's spillover ,and how similar it is to SPR. It's only used on targeted/constructed materials of independent sentences.

So we want to do it on some longer natural text and look for these underlying patterns. & see how well participants can read and understand longer texts.

The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers. At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions. Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants. If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts. The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.

We tweak the A-maze task to accomodate these longer passages and then have participants read them and also compare with  @futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to.

Using this tweak, we test A-maze on the passages from the Natural Stories corpus.  Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles show a similar pattern to SPR but with less noise and spillover

@@

```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
```
*mention the whole terminate on errors thing*
A-maze is a potentially powerful addition to the psycholinguists toolkit. However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences. This limits its useful, as some important questions such as comparing human data to language models or studying discourse effects require processing times over multi-sentence passages. Therefore, it's important to expand the Maze task to these types of materials and verify that it finds comparable patterns to other methods.

While the issue of needing to generate distractors for long passages is solved with A-maze, another problem with Maze remains. In particular, because Maze tasks discontinue after participants make mistakes, the farther into an item a word is, the fewer participants see it. This makes it hard to run long materials using Maze, and prevents Maze from being used on long text passages where SPR or eye-tracking could be used. We resolve this issue by introducing a new paradigm for running Maze tasks where participants correct their errors and continue reading rather than discontinuing after a mistake.

One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error. Thus, only participants who have processed the sentence and are paying attention contribute data at critical regions later in the sentence. However, this means we don't have data after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue. However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem. Well-chosen distractors and attentive participants will reduce the error rate, but the error rate will still compound over long materials. For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete a 50 word vignette, and `r round(.99**200*100)`% a 200 word passages.  In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.

@@

This variant of Maze expands the types of materials that can be used with maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.

@@

## Materials
We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], split them into sentences, and ran the sentences through the A-maze generation process.
We follow the A-maze generation process outlined in @boyceMazeMadeEasy2020, although we use an updated version of the codebase with fixes some issues identified in that paper (code at https://github.com/vboyce/Maze). Additionally, the capability to appropriately handle a wider variety of punctuation (needed for this corpus) was added. We took the auto-generated distractors as they were, without checking them for quality. 
We used the original comprehension questions provided in the Natural Stories corpus. To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions. See the Appendix for an except of one of the stories and its corresponding comprehension questions. All materials are available at TODO NEW REPO.

## Participants
We recruited 100 participants from Amazon Mechanical Turk, and paid each participant $3.50 for roughly 20 minutes of work. We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.

## Procedure
Participants first gave their informed consent and saw task instructions. Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task. After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter for payment. The experiment was implemented in Ibex (https://github.com/addrummond/ibex) and the experimental code is available at REPO.

@@

We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.

As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.

@@

```
To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].

We exclude the first word of every sentence because it has a different distractor, which leaves `r nrow(not_first)` words. Then we exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words. We additionally exclude words that were treated as being composed of multiple tokens. This is primarily words that have punctuation, where the punctuation is treated as a separate token. While surprisals should be additive, the summing the surprisals over these parts gives unreasonable responses sometimes. For instance, in one story the word king!' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, 'What is the use of flying any higher? This victory is in the bag and I am king!'). To avoid these outliers we exclude all words that a model treated as multi-token, leaving `r nrow(not_gpt)` words. (While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)

For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word. We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted). We exclude words where mistakes occurred or which occurred after a mistake in the same sentence. We only analysed words where we had values for all predictors, which means that if the previous word was unknown to a model, this word will be excluded because we're missing values for a lagged predictor.

For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability. We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions. To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word.

@@

For model comparison, we took by-item averaged data to aid in fast model fitting.  We included frequency, length, and their interaction to all models. Then we fit models with either 1 or 2 sources of surprisal using lm and assessed the effect of adding the second surprisal source with an anova. We used predictors for the current and past word and centered all effects.

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011. They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020]. For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
For linear models, we centered all predictors. We were unable to fit the full mixed effects model. The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.

@@

Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than other paradigms. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

We use task performance as our exclusion metric because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers). For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the questions correct.

@@

We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word. For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.

(ref:gam-cap) GAM predictions of reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom). Density of data is shown along the x-axis. For each of the 4 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data). There is no relationship between previous word surprisal and RT.

@@

The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam). Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal. All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data. All of the models show a flat relationship between previous word surprisal and RT. This is a sign of localization as the previous word's surprisal is not affecting RT, only the word's own suprisal is. The linear relationship matches that found with other methodologies.

@@

As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency. These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method. The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019. Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.

As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison shown in Table \@ref(tab:maze-compare). We assess the benefits of adding each model's predictions as a second surprisal source to see which models pick up on information not contained in another model. In terms of log likelihoods, GPT-2 is better than GRNN is better than TXL is better than 5-gram. Model comparisons tell a similar story; GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2. TXL provides a lot over 5-gram, and 5-gram provides little over any model. 
```{r maze-compare}

read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests are reported. We also report log likelihoods of models with only one surprisal source.")

@@

Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare). In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020. This contrasts with the Maze results, where the 5-gram model has the worst fit and does not provide additional predictive value to the other models.

@@

We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials. In this error-correction Maze paradigm, participants read all the words because they can correct their mistakes and move on. We tested this method on the Natural Stories corpus, showing that, despite the oddness of the task, participants can read and understand a 1000 word story in this method. We additionally showed that the RTs generated in the Maze task show a linear relationship between the RT of a word and it's surprisal, but no relationship with the surprisal of the previous word. This provides additional evidence for the argument that the Maze task forces very incremental processing [@forsterMazeTaskMeasuring2009].

>>>>

TODO overall: go through Roger's comments!
TODO fix tense!

@@

Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT. The two most commonly used methods are eye-tracking and self-paced reading, but both of these suffer from a lack of localization. In eye-tracking, participants read a text on a screen naturally, while their eye-movements are recorded [@raynerEyeMovementsReading1998]. The downside of this natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982]. Self-paced reading (SPR) is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead. Even in this method, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006]. Both these methods are prone to spillover, as the time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed, so by the time any late processing difficulties occur, the reader may already be several words along.

@@

## Frequency, length, and surprisal effects
One way of assessing how much a method localized RT effects is to look at how strongly properties of a word correlate with the RT of that word compared with RTs of downstream words later in the sentence. A couple of properties known to influence reading time are a word's length and overall frequency in a language, as longer and less common words take longer to process [@klieglLengthFrequencyPredictability2004]. Another measure is predictability, or how expected a word is given the context, with more predictable words being read faster [@raynerEffectsFrequencyPredictability2004]. A word can have low predictability for a number of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse. Many targeted effects of interest are essentially looking at specific features that contribute to how predictable or unpredictable a word is. This means that how good a method is at detecting and localizing effects of predictability is a key aspect of how useful it is as measure of incremental processing.
With al these factors potentially contributing to predictability, how is predictability as a whole measured? The typical method is to use language models that are trained on large corpora of language to predict what word comes next in a sentence. A variety of pre-trained models exist, that vary in their internals, but all of them will generate assessments of how predictable words are. Predictability is often represented in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc). 
The functional form of the relationship between RTs from eye-tracking and SPR corpora and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016]. Due to spillover effects, this linear relationship also holds between the surprisal of a previous word and the RT on the current word.

## Current experiment
The Maze task has thus far only been used on constructed sentences focusing on targeted effects and not on the sorts of long naturalistic passages used to assess the relationship between RT and surprisal. We want to test whether participants can read and understand stories while doing the demanding Maze task and also whether the RT profiles from the Maze task are similar to those from other methods.

The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers. At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions. Taken together, these properties are a severe test of A-maze, as these features may make it harder to auto-generate appropriate distractors and require focus from participants. If participants can succeed at the Maze task on Natural Stories, we think they are likely to succeed on a wide variety of naturalistic texts. The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.

We tweak the A-maze task to accomodate these longer passages and then have participants read the passages in the Maze. We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020. We find that participants were able to read and understand these long passages using A-maze, and their RT profiles show a similar pattern to SPR but with less noise and spillover

@@

```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
```

One of the benefits of the Maze task is that it forces incremental processing by having participants make an active choice about what the next word is. But what if they choose incorrectly? In the traditional Maze paradigm, there isn't a way to handle this; a mistake indicates they're not understanding the sentence properly, so the sentence ends, and the participant moves on to the next item [@forsterMazeTaskMeasuring2009]. An advantage of this is that participants who contribute RT data are very likely to have understand the sentence up to that point (it's hard to choose correctly in G-maze otherwise). This contrasts with other methods, where determining whether participants are paying attention usually requires separate comprehension check questions, that are usually not used for Maze.

However, terminating sentences on errors means that we don't have data after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue. However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem with long items. Well-chosen distractors and attentive participants will reduce the error rate, but the error rate will still compound over long materials. For instance, with a 1% error rate, `r round(.99**15*100)`%  of participants would complete each 15-word sentence, but only `r round(.99**50*100)`% would complete a 50 word vignette, and `r round(.99**200*100)`% would complete a 200 word passage.  In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.

@@

This variant of Maze expands the types of materials that can be used with Maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors. Error-correction Maze is a change in experimental procedure, and is independent of what types of distractors are used. This error-correction presentation is used here with A-maze, but would also work with G-maze or I-maze.

@@

## Materials
We used the texts from the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and generated A-maze distractors for them. We used the original comprehension questions provided in the Natural Stories corpus. To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions. See the Appendix for an except of one of the stories and its corresponding comprehension questions. All materials are available at TODO NEW REPO.

To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process. We used an updated version of the codebase from @boyceMazeMadeEasy2020. This newer version had the capability to match the greater variety of punctuation present in this corpus (updated auto-generation code at https://github.com/vboyce/Maze).  We took the auto-generated distractors as they were, without checking them for quality.

## Participants
We recruited 100 participants from Amazon Mechanical Turk in April 2020, and paid each participant $3.50 for roughly 20 minutes of work. We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.

## Procedure
Participants first gave their informed consent and saw task instructions. Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task. After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, and were debriefed, before getting a code to enter for payment. The experiment was implemented in Ibex (https://github.com/addrummond/ibex) and the experimental code is available at REPO.

@@

TODO top level description

In order to measure the relationship between a word's properties and it's RT, we fit models using surprisal, frequency, and length as predictors of RT. We considered these predictors from both the current and past word to look for the possibility of spill over effects. We fit generalized additive models (GAMs) to look at the functional form of the surprisal-RT relationship and linear models (LMs) to look at the size of the effects. As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.

 TODO FIX ME We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.

@@

```
### Predictors 
We created a set of predictor variables of frequency, word length, and surprisals from 4 language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then rescaled it to be the log2 frequency of expected occurrences in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020]. These models cover a range of common architectures. We can infer robustness if models provide similar results, but also compare between models to look at how well each model fits human data. For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.

### Exclusions
We exclude the first word of every sentence because it had an x-x-x distractor, which leaves `r nrow(not_first)` words. We exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words. We additionally exclude words that any model treated as being composed of multiple tokens. While surprisals should be additive, the summing the surprisals over these tokens gives some unreasonable responses. For instance, in one story the word king!\' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher? This victory is in the bag and I am king!\'). To avoid these outliers we exclude all words that any model treated as multi-token, leaving `r nrow(not_gpt)` words. This primarily excludes words with punctuation. (While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)

We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted). We exclude words where mistakes occurred or which occurred after a mistake in the same sentence. We only analysed words where we had values for all predictors, which means that if the previous word was unknown to a model, this word will be excluded because we're missing values for a lagged predictor.

### Model specification
For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability. We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions. To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word. TODO edit for bootstrapping

@@

For model comparison, we took by-item averaged data to aid in fast model fitting.  We included frequency, length, and their interaction to all models. Then we fit models with either 1 or 2 sources of surprisal using lm and assessed the effect of adding the second surprisal source with an anova. We used predictors for the current and past word and centered all effects. TODO cite LM

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011. They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].

For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
For linear models, we centered all predictors. We were unable to fit the full mixed effects model. The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.

@@

Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than in eye-tracking or SPR. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

We use task performance as our exclusion metric for A-maze because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers). For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the questions correct.

@@

TODO remind why we care so much about surprisal

We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word. For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded. (See Methods for details on exclusions.)

(ref:gam-cap) GAM predictions of reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom). Density of data is shown along the x-axis. For each of the 4 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data). The relationship between previous word surprisal and RT is much flatter.

@@

The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam). Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal. All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data. All of the models show a flatter relationship between previous word surprisal and RT. This is a sign of localization as the previous word's surprisal is not affecting RT much. The linear relationship matches that found with other methodologies.

@@

As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency. The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019. Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.

As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison shown in Table \@ref(tab:maze-compare). We assess the benefits of adding each model's predictions as a second surprisal source to see which models pick up on information not contained in another model.  GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2. TXL provides a lot over 5-gram, and 5-gram provides little over any model. Log likelihood, which is a measure of fit to data, shows  a similar hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
```{r maze-compare}

read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data. Each row shows the additional predictive value gained from adding that model to another model. F values and p values from anova tests between 1-surprisal-source and 2-source models are reported. We also report log likelihoods of models with only one surprisal source.")

@@

Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).  Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020.  This contrasts with the Maze results, where the 5-gram model has the worst fit and does not provide additional predictive value to the other models.

@@

## maze works

## strong effects of suprisal, mostly localized

## better than SPR

## correlated w SPR, probably the same effects, just stronger & less noisy

## Why you care - broad applications that might want this method

## shill for error correction
We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials. In this error-correction Maze paradigm, participants read all the words because they can correct their mistakes and move on. We tested this method on the Natural Stories corpus, showing that, despite the oddness of the task, participants can read and understand a 1000 word story in this method. We additionally showed that the RTs generated in the Maze task show a linear relationship between the RT of a word and it's surprisal, but no relationship with the surprisal of the previous word. This provides additional evidence for the argument that the Maze task forces very incremental processing [@forsterMazeTaskMeasuring2009].
